The Cost of Exclusion: Political Discrimination and Trust Deficits in the European Union

### 2025-07-14 article/main.tex
diagram updated

\begin{abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Integer tempus sagittis lobortis. Ut ac ligula a dolor faucibus rutrum. Donec accumsan sem nibh, quis ornare lorem volutpat a. Sed nec efficitur diam. Vivamus molestie turpis lorem, vitae semper lacus convallis sodales. Cras tellus leo, ultrices sit amet fringilla et, aliquam commodo tellus. Maecenas varius tincidunt libero, in aliquam orci iaculis eu. Vestibulum consequat ultrices eros id feugiat. Aliquam scelerisque porta lorem, id venenatis justo. Curabitur feugiat ante nec tellus pulvinar elementum. Orci varius natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Sed pellentesque ac dui id accumsan. Nam tellus libero, venenatis nec libero quis, varius rutrum velit. Quisque eu finibus diam. Nunc sit amet massa purus.
\end{abstract}

@@

While Europe has historically exhibited lower levels of affective polarization and declining partisan identification, recent studies suggest the presence of partisan hostility and negative feelings driven by factors such as ideological distance, perceived party threats, and regional tensions. Yet, evidence on longitudinal trends is still inconclusive. Scholars continue to emphasize the need for improved measurements and long-term data to determine whether affective polarization is indeed increasing across European democracies \parencite{wagner_affective_2024}.

>>>>

\begin{abstract}
This study examines whether experiences of political discrimination erode trust in public institutions across the European Union. Drawing on novel survey data from the EUROVOICES project—covering all 27 EU Member States and 110 subnational regions—we estimate the causal effect of discriminatory experiences using matching methods and sensitivity analyses. Findings show that individuals who report D/H experiences exhibit 16\% to 19\% lower trust in institutions. The effect is stronger among those who are politically unaligned or feel ideologically distant from the regional mainstream. By centering political discrimination—a dimension often overlooked in trust research—this study highlights how exclusion based on political beliefs can undermine democratic legitimacy and intensify public disaffection in polarized societies.
\end{abstract}

@@

While Europe has historically exhibited lower levels of affective and ideological polarization, recent studies suggest the rising presence of partisan hostility and negative feelings driven by factors such as ideological distance, perceived party threats, and regional tensions. Yet, evidence on longitudinal trends is still inconclusive. Scholars continue to emphasize the need for improved measurements and long-term data to determine whether polarization is indeed increasing across European democracies \parencite{wagner_affective_2024}.
